*Prochlorococcus marinus* responses to light and oxygen

### 2024-02-21 Code/Manuscript_O2.Rmd
Corrected spelling

To estimate the Photosynthetic Usable Radiation (PUR), a proxy of incident photons absorbed by the cells, for each *P. marinus* ecotype, the imposed Photosynthetic Active
Radiation (PAR) was first determined using the reported delivery of sinuisoidal diel PAR regimes by the Multicultivators, point validated using an XXXXXcitation of PAR meterXXX. An emission profile from 400 nm to 700 nm of each colored LED
light of the MCMIX-OD Multicultivator and the white LED light of the MC1000-OD Multicultivator was obtained using a Jaz spectrometer (Ocean Optics, Inc.,Dunedin, FL, USA) equipped with a fiber
optic cable, HH2 FiberOpticJmp (Part number A901073, Malvern Panalytical
Ltd, Malvern, UK). Each LED spectrum was then normalized to its emission maximum.
An *in-vivo* whole cell absorbance spectrum for each *P. marinus* strain under each spectral growth condition was obtained using the Olis 14 UV/VIS Clarity
Spectrophotometer (Olis Inc, Bogart, GA, USA) to scan across range of λ = 350 nm to 750 nm at 1 nm intervals. The path length of the internally reflective cavity of the Olis spectrophotometer was corrected to a 1 cm path length using the Javorfi correction method [@javorfiQuantitativeSpectrophotometryUsing2006] on PRO 99 media subtracted whole cell absorbance spectra. The blank-corrected whole cell absorbance spectra were normalized to the absorbance maximum of divinylchlorophyll a (Chl a~2~), determined for each spectra, falling between 400 nm and 460 nm.

@@

### MED4 Clade HLI  {.unnumbered}
The GAM model in Figure \@ref(fig:MED4GAM) shows MED4 growth in response to red (A) or blue (B) peak PAR and photoperiod across 2 imposed oxygen concentrations.  Under 250 µM O~2~ MED4 achieved fastest growth rates above peak blue light of ~180 µmol photons m^-2^ s^-1^, and the longest photoperiod of 16, indicated by the 0.69 d^-1^ contour line representing the
90^th^ percentile of maximum achieved growth rate.  Growth decreased with decreasing photoperiod and peak PAR.  Under red light growth was generally slower but the pattern of growth responses to photoperiod and PAR waas similar.  We again note the exclusion of MED4 from growth under 4 h photoperiod. Under 25 µM O~2~ MED4 showed similar growth responses, but was excluded from both 4 and 8 h photoperiods.

@@

Similar to the growth trends under the 250 µM O~2~ and PAR of 30 µmol
photons m^-2^ s^-1^ experiments, SS120 maintained at 25 µM O~2~ showed no
growthunder a 4 h photoperiod for imposed red spectral
waveband, however under blue light, SS120 was able to grow (Figure
\@ref(fig:SS120PAR)). In contrast to the growth trends of the 250 µM
O~2~ and PAR of 90 µmol photons m^-2^ s^-1^ experiments, SS120 grew
under 4 and 8 h blue light and 16 h red light photoperiods. The growth
rate decreased with increasing photoperiod under PAR of 90 µmol photons
m^-2^ s^-1^ white LED light treatments. Blue light treatments under PAR
of 180 µmol photons m^-2^ s^-1^ showed growth only under an 8 h
photoperiod. The µ~max~ was
`r PURPARFits %>% filter(Strain == "SS120", WL == "450", O2_uM == "25", Photoperiod == 8) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ (Table. \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons
m^-2^ s^-1^ blue light and 8 h photoperiod.

@@

The GAM model in Figure \@ref(fig:SS120GAM) shows growth
responses of SS120 red (A) or blue (B) peak PAR and photoperiod,
across the 2 imposed oxygen concentrations. Under 250 µM O~2~,
Figure \@ref(fig:SS120GAM)A showed highest growth
rates achieved below PAR of 50 µmol photons m^-2^ s^-1^ and
photoperiods between 8 and 12 h, indicated by the contour line labeled
0.29 d^-1^ (representing the 90^th^ percentile of achieved growth rate). Note the disjunct regions of the plots, resulting from variable growth successes of SS120 under 250 µM O~2~.  Growth rate patterns under red light were similar, although somehwat faster.
In contrast, under 25 µM O~2~ and a photoperiod of 8 h SS120 exploited all blue peak PAR levels, achieving faster growth ratesat a higher PAR of \~100 µmol photons
m^-2^ s^-1^, indicated by the contour line labeled 0.5 d^-1^
(representing the 90^th^ percentile of achieved growth rate), out pacing
the 90^th^ percentile fastest growth rates under 250 µM O~2~. Under red light and 25 µM O~2~ light SS120 grew across most conditions of peak PAR and photoperiod, achieving fastest growhth under long photoperiods and peak PAR below 50 \~100 µmol photons
m^-2^ s^-1^.  Thus, the designation of SS120 as a LL strain is dependent upon the [O~2]. SS120 did not, however,  grow under tested conditions at 2.5 µM O~2~.

@@

AllStrainPurFitsPlots_cap <- glue("**Chlorophyll specific growth rate (d^-1^) vs. cumulative diel Photosynthetic Usable Radiation (PUR, µmol photons m^-2^ d^-1^).**",
                    "Rows show levels of imposed dissolved O~2~ concentrations as 250 µM, 25 µM and 2.5 µM.",
                    "Columns are strains; MED4, SS120 and MIT9313.",
                    "Shapes show the imposed photoperiod (h); 4 h (hollow inverted triangle),  8 h (solid diamond), 12 h (solid circle), 16 h (solid upright triangle).",
                      "Symbol colours show the spectral waveband for growth; white LED (black symbols), 660 nm (red symbols), and 450 nm (blue symbols).",
                      "Large symbols show mean of growth rate from logistic curve fits; small symbols show values for replicate determinations, if any.",
                    "Platt (add Platt citation here) 4 parameter model fit to data pooled for each combination of strain and dissolved oxygen shown with solid lines. Seperate models fit to photoperiod data and shown if significantly different from the pooled model; 4 h (long dashed line); 8 h (dotted line); 12 h (dashed line); and 16 h (dot dashed line).",
                       .sep = " ")

@@

To gain insights into the variations we observed in growth responses across strains, particularly the responses to different oxygen levels [O~2~], we used a subset of the dataset of annotated phytoplankton genomes from @omarAnnotationGenesEncoding2023. This subset contains annotated phytoplankton gene sequences, with a focus on genes encoding enzymes involved in oxygen metabolism (Figure \@ref(fig:ProchlorococcusEnzymeKms)) or DNA repair (Figure \@ref(fig:DNARepairFig)). According to figure \@ref(fig:ProchlorococcusEnzymeKms), we see that *P. marinus* MED4 has the gene encoding ubiquinol oxidase (non electrogenic) enzyme, whereas SS120 and MIT9313 do not. The ubiquinol oxidase (non electrogenic) enzyme is involved in dissipating excess electrons [@InformationEC10] generated under high light exposure, thereby reducing stress on the electron transport system. The O~2~ substrate K~m~ value of approximately 25 µM  for ubiquinol oxidase  may explain why *P. marinus* MED4 can grow under 25 µM of O~2~ but not under 2.5 µM of O~2~, since growth under 2.5 µM of O~2~ would suppress use of ubiquinol oxidase to balance electron transport.  The demonstrated ability of *P. marinus* MIT9313 to grow when exposed to high blue light under 2.5 µM and 25 µM O~2~ treatments may be explained by the presence of the gene encoding enzyme 2-hydroxy-acid oxidase (Figure \@ref(fig:ProchlorococcusEnzymeKms)). The enzyme 2-hydroxy-acid oxidase, with an O~2~ substrate K~m~ value of approximately 250 µM (Figure \@ref(fig:ProchlorococcusEnzymeKms)), catalyzes the reaction of 2-hydroxy acid with O~2~ to produce toxic hydrogen peroxide [@InformationEC15]. We hypothesize that under our 250 µM O~2~ and high blue light experiments, *P. marinus* MIT9313 exhibited photoinhibition, resulting from the inactivation of PSII caused by the production of the reactive oxygen species, hydrogen peroxide. This inhibition was further compounded by the limited induced repair mechanism for PSII, due to the absence of FtsH 2 and 3  proteins in *P. marinus* MIT9313 [@bonisteelStrainSpecificDifferences2018].  We hypothesize that under the conditions of our high light and 2.5 µM and 25 µM O~2~ experiments, the activity of the 2-hydroxy-acid oxidase enzyme is suppressed. As a result, the catalyzed production of hydrogen peroxide is inhibited, leading to minimal PSII damage, allowing MIT9313 to avoid photohibition and circumvent limitations on PSII repair.  Figure \@ref(fig:ProchlorococcusEnzymeKms) also shows that *P. marinus* SS120 is the only tested ecotype to lack the pyridoxal 5'-phosphate synthase enzyme. The pyridoxal 5'-phosphate synthase enzyme is an important cofactor in the biosynthesis of vitamin B~6~ [@sangIdentificationPyridoxinePyridoxamine2007]. Vitamin B~6~ is a potential antioxidant and can effectively quench singlet oxygen [@bilskiVitaminB6Pyridoxine2000]. The absence of the pyridoxal 5'-phosphate synthase enzyme may explain why *P. marinus* SS120 does not grow as well as *P. marinus* MIT9313 when exposed to high light stress under 25 µM O~2~ and not at all under 2.5 µM O~2~.

>>>>

To estimate the Photosynthetic Usable Radiation (PUR), a proxy of incident photons absorbed by the cells, for each *P. marinus* ecotype, the imposed Photosynthetic Active
Radiation (PAR) was first determined using the reported delivery of sinusoidal diel PAR regimes by the Multicultivators, point validated using an XXXXXcitation of PAR meterXXX. An emission profile from 400 nm to 700 nm of each colored LED
light of the MCMIX-OD Multicultivator and the white LED light of the MC1000-OD Multicultivator was obtained using a Jaz spectrometer (Ocean Optics, Inc.,Dunedin, FL, USA) equipped with a fiber
optic cable, HH2 FiberOpticJmp (Part number A901073, Malvern Panalytical
Ltd, Malvern, UK). Each LED spectrum was then normalized to its emission maximum.
An *in-vivo* whole cell absorbance spectrum for each *P. marinus* strain under each spectral growth condition was obtained using the Olis 14 UV/VIS Clarity
Spectrophotometer (Olis Inc, Bogart, GA, USA) to scan across range of λ = 350 nm to 750 nm at 1 nm intervals. The path length of the internally reflective cavity of the Olis spectrophotometer was corrected to a 1 cm path length using the Javorfi correction method [@javorfiQuantitativeSpectrophotometryUsing2006] on PRO 99 media subtracted whole cell absorbance spectra. The blank-corrected whole cell absorbance spectra were normalized to the absorbance maximum of divinylchlorophyll a (Chl a~2~), determined for each spectra, falling between 400 nm and 460 nm.

@@

### MED4 Clade HLI  {.unnumbered}
The GAM model in Figure \@ref(fig:MED4GAM) shows MED4 growth in response to red (A) or blue (B) peak PAR and photoperiod across 2 imposed oxygen concentrations.  Under 250 µM O~2~ MED4 achieved fastest growth rates above peak blue light of ~180 µmol photons m^-2^ s^-1^, and the longest photoperiod of 16, indicated by the 0.69 d^-1^ contour line representing the
90^th^ percentile of maximum achieved growth rate.  Growth decreased with decreasing photoperiod and peak PAR.  Under red light growth was generally slower but the pattern of growth responses to photoperiod and PAR was similar.  We again note the exclusion of MED4 from growth under 4 h photoperiod. Under 25 µM O~2~ MED4 showed similar growth responses, but was excluded from both 4 and 8 h photoperiods.

@@

Similar to the growth trends under the 250 µM O~2~ and PAR of 30 µmol
photons m^-2^ s^-1^ experiments, SS120 maintained at 25 µM O~2~ showed no
growth under a 4 h photoperiod for imposed red spectral
waveband, however under blue light, SS120 was able to grow (Figure
\@ref(fig:SS120PAR)). In contrast to the growth trends of the 250 µM
O~2~ and PAR of 90 µmol photons m^-2^ s^-1^ experiments, SS120 grew
under 4 and 8 h blue light and 16 h red light photoperiods. The growth
rate decreased with increasing photoperiod under PAR of 90 µmol photons
m^-2^ s^-1^ white LED light treatments. Blue light treatments under PAR
of 180 µmol photons m^-2^ s^-1^ showed growth only under an 8 h
photoperiod. The µ~max~ was
`r PURPARFits %>% filter(Strain == "SS120", WL == "450", O2_uM == "25", Photoperiod == 8) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ (Table. \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons
m^-2^ s^-1^ blue light and 8 h photoperiod.

@@

The GAM model in Figure \@ref(fig:SS120GAM) shows growth
responses of SS120 red (A) or blue (B) peak PAR and photoperiod,
across the 2 imposed oxygen concentrations. Under 250 µM O~2~,
Figure \@ref(fig:SS120GAM)A showed highest growth
rates achieved below PAR of 50 µmol photons m^-2^ s^-1^ and
photoperiods between 8 and 12 h, indicated by the contour line labeled
0.29 d^-1^ (representing the 90^th^ percentile of achieved growth rate). Note the disjunct regions of the plots, resulting from variable growth successes of SS120 under 250 µM O~2~.  Growth rate patterns under red light were similar, although somewhat faster.
In contrast, under 25 µM O~2~ and a photoperiod of 8 h SS120 exploited all blue peak PAR levels, achieving faster growth rates at a higher PAR of \~100 µmol photons
m^-2^ s^-1^, indicated by the contour line labeled 0.5 d^-1^
(representing the 90^th^ percentile of achieved growth rate), out pacing
the 90^th^ percentile fastest growth rates under 250 µM O~2~. Under red light and 25 µM O~2~ light SS120 grew across most conditions of peak PAR and photoperiod, achieving fastest growth under long photoperiods and peak PAR below 50 \~100 µmol photons
m^-2^ s^-1^.  Thus, the designation of SS120 as a LL strain is dependent upon the [O~2]. SS120 did not, however,  grow under tested conditions at 2.5 µM O~2~.

@@

AllStrainPurFitsPlots_cap <- glue("**Chlorophyll specific growth rate (d^-1^) vs. cumulative diel Photosynthetic Usable Radiation (PUR, µmol photons m^-2^ d^-1^).**",
                    "Rows show levels of imposed dissolved O~2~ concentrations as 250 µM, 25 µM and 2.5 µM.",
                    "Columns are strains; MED4, SS120 and MIT9313.",
                    "Shapes show the imposed photoperiod (h); 4 h (hollow inverted triangle),  8 h (solid diamond), 12 h (solid circle), 16 h (solid upright triangle).",
                      "Symbol colours show the spectral waveband for growth; white LED (black symbols), 660 nm (red symbols), and 450 nm (blue symbols).",
                      "Large symbols show mean of growth rate from logistic curve fits; small symbols show values for replicate determinations, if any.",
                    "Platt (add Platt citation here) 4 parameter model fit to data pooled for each combination of strain and dissolved oxygen shown with solid lines. Separate models fit to photoperiod data and shown if significantly different from the pooled model; 4 h (long dashed line); 8 h (dotted line); 12 h (dashed line); and 16 h (dot dashed line).",
                       .sep = " ")

@@

To gain insights into the variations we observed in growth responses across strains, particularly the responses to different oxygen levels [O~2~], we used a subset of the dataset of annotated phytoplankton genomes from @omarAnnotationGenesEncoding2023. This subset contains annotated phytoplankton gene sequences, with a focus on genes encoding enzymes involved in oxygen metabolism (Figure \@ref(fig:ProchlorococcusEnzymeKms)) or DNA repair (Figure \@ref(fig:DNARepairFig)). According to figure \@ref(fig:ProchlorococcusEnzymeKms), we see that *P. marinus* MED4 has the gene encoding ubiquinol oxidase (non electrogenic) enzyme, whereas SS120 and MIT9313 do not. The ubiquinol oxidase (non electrogenic) enzyme is involved in dissipating excess electrons [@InformationEC10] generated under high light exposure, thereby reducing stress on the electron transport system. The O~2~ substrate K~m~ value of approximately 25 µM  for ubiquinol oxidase  may explain why *P. marinus* MED4 can grow under 25 µM of O~2~ but not under 2.5 µM of O~2~, since growth under 2.5 µM of O~2~ would suppress use of ubiquinol oxidase to balance electron transport.  The demonstrated ability of *P. marinus* MIT9313 to grow when exposed to high blue light under 2.5 µM and 25 µM O~2~ treatments may be explained by the presence of the gene encoding enzyme 2-hydroxy-acid oxidase (Figure \@ref(fig:ProchlorococcusEnzymeKms)). The enzyme 2-hydroxy-acid oxidase, with an O~2~ substrate K~m~ value of approximately 250 µM (Figure \@ref(fig:ProchlorococcusEnzymeKms)), catalyzes the reaction of 2-hydroxy acid with O~2~ to produce toxic hydrogen peroxide [@InformationEC15]. We hypothesize that under our 250 µM O~2~ and high blue light experiments, *P. marinus* MIT9313 exhibited photoinhibition, resulting from the inactivation of PSII caused by the production of the reactive oxygen species, hydrogen peroxide. This inhibition was further compounded by the limited induced repair mechanism for PSII, due to the absence of FtsH 2 and 3  proteins in *P. marinus* MIT9313 [@bonisteelStrainSpecificDifferences2018].  We hypothesize that under the conditions of our high light and 2.5 µM and 25 µM O~2~ experiments, the activity of the 2-hydroxy-acid oxidase enzyme is suppressed. As a result, the catalyzed production of hydrogen peroxide is inhibited, leading to minimal PSII damage, allowing MIT9313 to avoid photoinhibition and circumvent limitations on PSII repair.  Figure \@ref(fig:ProchlorococcusEnzymeKms) also shows that *P. marinus* SS120 is the only tested ecotype to lack the pyridoxal 5'-phosphate synthase enzyme. The pyridoxal 5'-phosphate synthase enzyme is an important cofactor in the biosynthesis of vitamin B~6~ [@sangIdentificationPyridoxinePyridoxamine2007]. Vitamin B~6~ is a potential antioxidant and can effectively quench singlet oxygen [@bilskiVitaminB6Pyridoxine2000]. The absence of the pyridoxal 5'-phosphate synthase enzyme may explain why *P. marinus* SS120 does not grow as well as *P. marinus* MIT9313 when exposed to high light stress under 25 µM O~2~ and not at all under 2.5 µM O~2~.
